Natural sources of vagueness and their implications

### 2017-05-10 book_chapter/paper.tex
typos, minor polishing and some inline comments on more juicy matters

\begin{abstract}
A vexing puzzle about vagueness, rationality and evolution runs, in crude abbreviation, as follows: vague language use is demonstrably suboptimal if the goal is efficient precise and cooperative information transmission; hence rational deliberation or evolutionary selection should, under this assumed goal, eradicate vagueness from language use.
Since vagueness is persistent in all human languages, something has to give.
In this paper, we investigate a number of reasons why and mechanisms how vagueness may come into the picture in formal models of rational or evolutionary optimal signaling.
We show how uncertainty about not only the linguistic practices of others, but also about the world itself can lead to vagueness, and how, given vagueness, natural linguistic practices are likely to create more reason for vagueness.
We explore the consequences of these reasons and mechanism for a notion of meaning and a notion of language.
\end{abstract}

@@

The classic philosophical problem of vagueness is most starkly embodied by the sorites paradox.
The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, we can repeatedly remove all but one grain of sand of something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, we would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
Neither choice is, however, intuitively satisfying.
The paradox is interesting because it can be made general and re-applied to many other words besides `heap'.
Predicates for which one can find a suitable instance of the general formulation of the sorites paradox are called \emph{vague}.
Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\parencite{Keefe1997}.
How widespread is the problem?
It is easy to find more examples of predicates based on more finely grained properties---as `tall' is intuitively based on height---for which constructing a sorites paradox would be easy.
Mereological nihilists argue that instances of the sorites can be designed for any material object that can be decomposed into small enough parts.
If we subscribe to the scientific picture of matter as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
The problem thus seems to be a wide-ranging one, but it does not obviously undermine our everyday linguistic practices.
Therefore, the issue must lie with the conceptions we have of the forces or mechanisms that underlie those practices.

Vagueness is typically seen as a challenge to a classical conception of language that relies on reference and truth as two core notions for understanding meaning.
One problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to something like a well defined set of tall people?
And if we cannot know exactly whether a certain person is `tall' or not, as with borderline cases, how are we to determine the truth value of sentences that involve statements of tallness regarding these borderline cases?
Supervaluationism, many-valued logics, and degree theories all propose changes to the classical picture in order to accommodate for vague predicates.
They are, however, still strongly committed to its two core notions.
As argued by Mark Sainsbury~\parencite*{sainsbury_concepts_1999}, because of that they fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
All the aforementioned proposals end up being committed to new artificial demarcating boundaries (\emph{e.g.}~true-under-all-precisifications versus neither true nor false versus false-under-all-precisifications, true versus indefinite versus false, true to degree 1 versus true to degree 0 versus the rest).
But a vague predicate not only fails to demarcate between the cases where it clearly applies and the ones where it clearly doesn't, it also fails to establish a boundary between the cases where it clearly applies and the borderline cases, as well as between the borderline cases and the cases where it clearly doesn't apply.
Further introducing borderline borderline cases would lead to an infinite regress.
Because of their attachment to the classical picture, the standard approaches to vagueness also fail to draw an important lesson from the sorites paradox, namely that ``we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly''~\parencite*[256]{sainsbury_concepts_1999}.
% \begin{quote}
% But to what in our actual use of language does this division correspond?
% It looks as if, as before, it should correspond to the sentences true beyond the shadow of vagueness, those in some kind of borderline position, and those false beyond the shadow.
% But [\ldots] we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly.
% Hence they cannot be included in a correct description of our language.%
% ~\parencite[256]{sainsbury_concepts_1999}
% \end{quote}
By trying to cling as much as possible to the classical picture of logic and semantics, these standard approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~are able to recognize the logical inconsistency but do not have a good answer to overcome it.
Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language productively.
Nobody ever stopped using the word `tall` after being confronted with a sorites series.
Why should we develop theories of meaning that are impervious to the paradox?

The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also consequently need to be abandoned down the line.
One notion that seems to quickly be in peril is that of rationality.
In reference to philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
\begin{quote}
In the past, such philosophers have typically conjoined the claim that there is universal human agreement on the supreme desirability of truth with two further premises: that truth is correspondence to reality, and that reality has an intrinsic nature (that there is, in Nelson Goodman's terms, a Way the World Is).
[\ldots]
The rise of relatively democratic, relatively tolerant, societies in the last few hundred years is said to be due to the increased rationality of modern times, where `rationality' denotes the employment of an innate truth-oriented faculty.%
~\parencite*[1]{rorty_response_2000-1}
\end{quote}
By tying all of these notions tightly together, the fear could be that by dropping the picture of meaning as intimately tied to truth and logic we necessarily lose the ground on which rationality stands, and the whole edifice could collapse.
Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
Although humans are not necessarily consciously aware of the rules, they are considered to be nevertheless unconsciously following them when they act rationally.
This type of rationality is one that supposedly demarcates humans from other animals.
We want to argue here that giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
What we need is to revisit the notion of rationality tied to the classical picture.

Although frequently debated, especially in areas like philosophy, psychology, and economics, rationality is a notion for which clear definitions are hard to come by.
Discussions around it usually rely on common connotations and implicit assumptions.
The aforementioned classical picture of language and meaning is focused on an aspect of rationality that we can call \emph{procedural}.
It attempts to characterize a particular choice of an agent as rational based on the procedures used to select it.
We can also focus instead on the consequences, rather than the procedures.
\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if, in making that choice, the means available to him were employed in the best way possible that enables him achieve a certain goal.
%This is close to the notion of rationality that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility.
The notion is linked to David Hume~\parencite*{hume_treatise_1738} and his assertion that ``Reason is, and ought to only be the slave of the passions.''
Another assumption that is important to qualify when talking about rationality of an agent's choice is the amount of information that the agent has access to, when making his decision.
Depending on how accurate and complete the agent's knowledge of its environment, the objectives to be achieved, the choices available, the relation between those choices and the objectives, etc, the verdict over the rationality of his choices can vary.
This is difficult to characterize in general, but, in the context of a given model, we can describe agents with respect to how much knowledge they have of the ingredients of the model.
An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
Finally, the third assumption we will draw attention to, for now, regards the ability of an agent to take predictions into account
Thus we can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate and consider either the actions of other agents, or considerations of medium/long term gains.
The classical picture is, in these terms, tied to a procedural account of rationality, where agents are pretty much omniscient.
Questions of how strategic they are do not really come to the fore, since the analysis is typically both atemporal and individualistic.

What kind of rationality do we get if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
What could even such a paradigm be?
Perhaps we can draw inspiration from the later work of philosopher Ludwig Wittgenstein%
\footnote{To the extent that substantial views can be said to be defended by the author. Skipping over the debate (but see~\cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
Most starkly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953}, we can find a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.

Following this idea, we here survey proposals for addressing vagueness in the context of signaling games.
The following section introduces the framework of signaling games.
Then in sections~\ref{sec:sim-max-vagueness} and~\ref{sec:referential-vagueness} we discuss two particular types of games, namely similarity maximization and referential games, and which suggestions there are with regards to how vagueness can be accounted for in those models.
We bring the insights together in section~\ref{sec:finite-experience} and reflect on the consequences of the proposed natural sources of vagueness for notions of rationality.

@@

Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function%
\footnote{This could also be represented with two separate functions. In cases of perfectly aligned preferences, the function can be simplified to yield only one value, to be equally attributed to each player.}
$U : T \times A \rightarrow \mathbb{R}^2$ which yields a pair of payoff values, one for the sender, one for the receiver, for each possible pairing of state and action.

@@

The expected utility $EU$ of a pair of strategies $(\sigma,\rho)$ can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions. %, \emph{i.e.}:

@@

These would have an expected utility of $2$ for both sender and receiver, since when $t_1$ obtains the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, when $t_2$ obtains the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.

@@

The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like one would have in the classical picture.
However, we can already see how it attempts to address the study of meaning from a very different angle.
There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes.
There is also no appeal to truth as a guide for our inquiry.
One advantage of such a paradigm shift is that we are no longer tied to the need to explain how language hooks on to the world; this question is no longer relevant on this account.
We can merely focus on trying to see how agents can use signals to cope with the world and achieve their purposes.
On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
In the examples discussed so far, agents are assumed to possess and exercise rationality.
It is not, however, the procedural notion of rationality of the classical picture, but rather one that is instrumental and heavily strategic.

@@

In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes we study signaling problems in evolutionary terms.
Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
Although the inspiration for this scenario is biological evolution, similar things could be said about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}, \emph{i.e.}~we can interpret these notions in terms of cultural evolution~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985}.
The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
The only thing relevant to this equation are the relative proportions of strategies in a given population and the utility function.
Using it, one can compute which strategies evolve under which conditions.

Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalism to the descriptive side of the investigation.
Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
Focus is put on understanding how various ingredients of the model interact, and which results they produce, not on metaphysical concerns.
While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
This range of options goes hand in hand with a range of pictures of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of instrumental but limited and non-strategic rationality in the case of learning dynamics, to higher levels of rationality and even game-theoretical reasoning.
Each of these can be utilized depending on the problem that one is interested in characterizing.
Thus, although Skyrms shows that rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where rationality can play a role.

>>>>

\begin{abstract}
A vexing puzzle about vagueness, rationality and evolution runs, in crude abbreviation, as follows: vague language use is demonstrably suboptimal if the goal is efficient precise and cooperative information transmission; hence rational deliberation or evolutionary selection should, under this assumed goal, eradicate vagueness from language use.
Since vagueness is persistent in all human languages, something has to give.
In this paper, we investigate a number of reasons why and mechanisms how vagueness may come into the picture in formal models of rational or evolutionary optimal signaling.
We show how uncertainty about not only the linguistic practices of others, but also about the world itself can lead to vagueness, and how, given vagueness, natural linguistic practices are likely to create more reason for vagueness.
We explore the consequences of these reasons and mechanisms for a notion of meaning and a notion of language.
\end{abstract}

@@

The classical philosophical problem of vagueness is most starkly embodied by the sorites paradox.
The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, we can repeatedly remove all but one grain of sand of something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, we would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
Neither choice is, however, intuitively satisfying.
The paradox is interesting because it can be made general and re-applied to many other words besides `heap'.
Predicates for which one can find a suitable instance of the general formulation of the sorites paradox are called \emph{vague}.
Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\parencite{Keefe1997}.
How widespread is the problem?
It is easy to find more examples of predicates based on more finely grained properties---as `tall' is intuitively based on height---for which constructing a sorites paradox would be easy.
Mereological nihilists argue that instances of the sorites can be designed for any material object that can be decomposed into small enough parts.
If we subscribe to the scientific picture of matter as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
The problem thus seems to be a wide-ranging one, but it does not obviously undermine our everyday linguistic practices.
Most of the time we seem to communicate just fine.
Therefore, the issue must lie with the conceptions we have of the forces or mechanisms that underlie those practices.

Vagueness is typically seen as a challenge to a classical conception of language that relies on reference and truth as two core notions for understanding meaning.
One problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to something like a well defined set of tall people?
And if we cannot know exactly whether a certain person is `tall' or not, as with borderline cases, how are we to determine the truth value of sentences that involve statements of tallness regarding these borderline cases?
Supervaluationism, many-valued logics, and degree theories all propose changes to the classical picture in order to accommodate for vague predicates.
They are, however, still strongly committed to its two core notions.
As argued by Mark Sainsbury~\parencite*{sainsbury_concepts_1999}, because of that they fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
All the aforementioned proposals end up being committed to new artificial demarcating boundaries (\emph{e.g.}~true-under-all-precisifications versus neither true nor false versus false-under-all-precisifications, true versus indefinite versus false, true to degree 1 versus true to degree 0 versus the rest).
But a vague predicate not only fails to demarcate between the cases where it clearly applies and the ones where it clearly doesn't, it also fails to establish a boundary between the cases where it clearly applies and the borderline cases, as well as between the borderline cases and the cases where it clearly doesn't apply.
Further introducing borderline borderline cases would lead to an infinite regress.
Because of their attachment to the classical picture, the standard approaches to vagueness also fail to draw an important lesson from the sorites paradox, namely that ``we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly''~\parencite*[256]{sainsbury_concepts_1999}.
% \begin{quote}
% But to what in our actual use of language does this division correspond?
% It looks as if, as before, it should correspond to the sentences true beyond the shadow of vagueness, those in some kind of borderline position, and those false beyond the shadow.
% But [\ldots] we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly.
% Hence they cannot be included in a correct description of our language.%
% ~\parencite[256]{sainsbury_concepts_1999}
% \end{quote}
By trying to cling as much as possible to the classical picture of logic and semantics, these standard approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~they are able to recognize the logical inconsistency but do not have a good answer as to how to overcome it.
Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language productively.
Nobody ever stopped using the word `tall` after being confronted with a sorites series.
Why should we develop theories of meaning that are impervious to the paradox?

The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also need to be abandoned as a consequence.
One notion that seems to quickly be in peril is that of rationality.
In reference to philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
\begin{quote}
In the past, such philosophers have typically conjoined the claim that there is universal human agreement on the supreme desirability of truth with two further premises: that truth is correspondence to reality, and that reality has an intrinsic nature (that there is, in Nelson Goodman's terms, a Way the World Is).
[\ldots]
The rise of relatively democratic, relatively tolerant, societies in the last few hundred years is said to be due to the increased rationality of modern times, where `rationality' denotes the employment of an innate truth-oriented faculty.%
~\parencite*[1]{rorty_response_2000-1}
\end{quote}
By tying all of these notions tightly together, the fear could be that by dropping the picture of meaning as intimately tied to truth and logic we necessarily lose the ground on which rationality stands, and the whole edifice could collapse.
Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
Although humans are not necessarily consciously aware of the rules, they are considered to be nevertheless unconsciously following them when they act rationally.
This type of rationality is one that supposedly demarcates humans from other animals.
We want to argue here that giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore irrational.
What we need is to revisit the notion of rationality tied to the classical picture.

Although frequently debated, especially in areas like philosophy, psychology, and economics, rationality is a notion for which clear definitions are hard to come by.
Discussions around it usually rely on common connotations and implicit assumptions.
The aforementioned classical picture of language and meaning is focused on an aspect of rationality that we can call \emph{procedural}.
It attempts to characterize a particular choice of an agent as rational based on the procedures used to select it.
We can also focus instead on the consequences, rather than the procedures.
\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if, in making that choice, the available means were employed in the best way possible to achieve a certain goal.
%This is close to the notion of rationality that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility.
The notion is linked to David Hume~\parencite*{hume_treatise_1738} and his assertion that ``Reason is, and ought to only be the slave of the passions.''
Another assumption that is important to qualify when talking about rationality of an agent's choice is the amount of information that the agent has access to, when making his decision.
Depending on how accurate and complete the agent's knowledge of its environment, the objectives to be achieved, the choices available, the relation between those choices and the objectives, etc, the verdict over the rationality of his choices can vary.
This is difficult to characterize in general, but, in the context of a given model, we can describe agents with respect to how much knowledge they have of the ingredients of the model.
An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
Finally, the third assumption we will draw attention to, for now, regards the ability of an agent to take predictions into account.
Thus we can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate and consider either the actions of other agents, or considerations of medium/long term gains.
The classical picture is, in these terms, tied to a procedural account of rationality, where agents are pretty much omniscient.
Questions of how strategic they are do not really come to the fore, since the analysis is typically both atemporal and individualistic.
\mf{I suggest to rewrite this paragraph, keeping as much as possible, but to include a formal definition of ``instrumental rationality'' as maximization of expected utilities, and to also include notions ``ecological rationality'' and ``bounded rationality'' with special-purpose and clearly spelled out definitions, namely roughly like this: where ``instrumental rationality'' requires a single choice to maximize EU in a single local context / a single stage-game, \emph{ecological rationality} requires a general way of choosing (call it, e.g., a \emph{choice mechanism}) to be optimal (max EU) across a whole class of different contexts / games / decision-making situations; a choice mechanism is \emph{bounded rationality} if it is ecologically rational if we also take into account the processing costs necessary to reach a decision (so a choice mechanism that is fast-and-cheap can be better (under bounded rationality) than a highly precise one on average because it saves processing costs, while it would be strictly outperformed in any game/context/decision situation by a more precise mechanism)}

What implications for a notion of rationality of language users do we get if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
What could such a paradigm even be?
Perhaps we can draw inspiration from the later work of philosopher Ludwig Wittgenstein%
\footnote{To the extent that substantial views can be said to be defended by the author. Skipping over the debate (but see~\cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
Most starkly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953}, we can find a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.

Following this idea, we here survey proposals for addressing vagueness in the context of signaling games.
The following section introduces the framework of signaling games.
Section~\ref{sec:sim-max-vagueness} looks at explanations for vagueness in a special kind of signaling game, a so-called sim-max game.
Section~\ref{sec:finite-experience} reflects on the consequences of these explanations for notions of rationality.
Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game. 
\mf{Please check if this overview paragraph is appropriate or if you would like to go into a little more detail.}

@@

Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a probability measure $P$ such that $P(t)$ is the probability of frequency with which $t \in T$ occurs, a set of available signals or messages $M$, a set of responses or actions $A$, and a pair of utility functions $U_{S,R} : T \times A \rightarrow \mathbb{R}$, one for the sender and one for the receiver, each of which yields a payoff value for each possible pairing of state and action.

@@

The expected utility $EU$ of a strategy can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions, weighted by the probability of each state. Concretely, the expected utility of $\sigma$ given $\rho$ is $EU_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$ and the expected utility of $\rho$ given $\sigma$ is $EU_S(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
The expected u %, \emph{i.e.}:

@@

These would have an expected utility of $1$ for both sender and receiver, since when $t_1$ obtains with probability $P(t_1)$ the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, and when $t_2$ obtains with probability $P(t_2) = 1-P(t_1)$ the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.

@@

The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like one would have in the classical picture.
However, we can already see how it attempts to address the study of meaning from a very different angle.
There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes (but see \cite{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis} for characterizations of signal meaning in terms of some correspondence with states or acts).
There is also no appeal to truth as a guide for our inquiry.
One advantage of such a paradigm shift is that we are no longer tied to the need of explaining how language hooks on to the world; this question is no longer of primary relevance.
We can merely focus on trying to see how agents can use signals to cope with the world and achieve their purposes.
On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
In the examples discussed so far, agents are assumed to possess and exercise rationality.
It is not, however, the procedural notion of rationality of the classical picture, but rather one that is instrumental and heavily strategic.

@@

In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes to study signaling problems in evolutionary terms.
Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
Although the inspiration for this scenario is biological evolution, similar things could be said about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}, \emph{i.e.}~we can interpret these notions in terms of cultural evolution~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985}.
The principles can be captured in formal models that abstracts away from details of single interactions and individual agent's behavior, \emph{e.g.}~the replicator dynamics~\parencite{TaylorJonker1978:Evolutionary-St}.
The only thing relevant to the replicator dynamics are the relative proportions of strategies in a given population and the utility function.
Using it, one can compute which strategies evolve under which conditions.

Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalism to the descriptive side of the investigation.
Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
Focus is put on understanding how various ingredients of the model interact, and which results they produce, not on metaphysical concerns.
While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
This range of options goes hand in hand with a range of pictures of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of instrumental but limited and non-strategic rationality in the case of learning dynamics, to higher levels of rationality and even recursive strategic reasoning about the co-players' beliefs and choices.
Each of these can be utilized depending on the problem that one is interested in characterizing.
Thus, although Skyrms shows that rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where rationality can play a role.

The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under certain initial conditions and standard evolutionary pressures, and where it remains after small disturbances.
With these tools, not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which do not.
In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely. %; which one of the possible two will depend on their relative proportions in the original population.
More complex signaling problems may have different evolutionary outcomes, sometimes unexpected ones.
Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few.
More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \citet{FrankeWagner2014:Game-Theory-and}.
